---
title: "Module01"
author: "Caterina"
date: "2026-05-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Add a New Chunk 
option + command + i 

```{r}
#install.packages("tidyverse")
library(tidyverse)
library(rmarkdown)
```
#2.16 Load Packages and Data 

```{r}
#install.packages("knitr")
library(knitr)
mpg
```

```{r}
?mpg
```

# 2.18.1 Understanding the 'grammar of graphics' 
This is the template to build a graph... 

ggplot(data = <YOUR_DATA>) +
<GEOM_FUNCTION>(mapping = aes(<YOUR_MAPPINGS>)) 

# 2.17 Create the first ggplot2
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy))
```
# 2.18.2 Asthetics Mappings - making the graph pretty 
Now the points on the graph are colour coordinated depending on car class. 

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, colour = class))
```

BUT... instead of coordinating by class, we can do it by size... 

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, size = class))
```
This isn't a great looking graph, can't really see the sizes of the points. 
So, we are going to change the class to alpha and see what that looks like... 

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, alpha = class))
```
You also arent really able to see the difference between class of car. 
So, we try something different. We will try shape. 

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, shape = class))
```
This also doesn't work very well because the program can only do a maximum of 6 different shapes, and we have 7. 

Now we will ask R to make the points on the graph all blue. Remember, the code looks different, be careful. 

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy), colour = "blue")
```

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, colour = displ < 5))
```
HERE we keep the colour in the big brackets because we are changing information within the aesthetics, rather than the whole graph. 

# 2.19 Troubleshooting 
Be careful when you're typing code, the placement of specific numbers of letters etc. matter. E.g. 
```{r}
#ggplot(data = mpg) 
# + geom_point(mapping = aes(x = displ, y = hwy, colour = displ < 5))
```

The above doesn't work, because the + symbol is in the wrong place. 

# 2.20 Facet and Panel Plots 
This creates separate plots or subplots - these are called facets. 
We use the function, facet_wrap(). But, we can only use facet_wrap() for discrete variables. 

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_wrap(~class, nrow = 2)
```
We will now use facet_grid if we want to use more than one variable, using ~ to split them. 
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ cyl)
```
We will now only split into 1 variable. To do this, we use full stop. 
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, colour = drv)) +
  facet_grid(. ~ cyl)
```
Read facet_wrap using below code. 

```{r}
?facet_wrap
```

# 2.21 Fitting Simple Lines 
ggplot can use a variety of visual objects to represent our data. To display data as points, use geom_point(). 

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy))
```
HOWEVER< ggplot2 can use a variety of geom objects to represent the data. 
We might want to use bar plots, line charts, boxplots etc. We will now plot a smooth line. 

```{r}
ggplot(data = mpg) +
  #geom_point(mapping = aes(x = displ, y = hwy)) # points not as cute. 
  geom_smooth(mapping = aes(x = displ, y = hwy), colour = "red") #smooth line is nice. 
```
We can also change the line type. 

```{r}
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv), colour = "blue")
```
And then we can change the colour of the line by adding colour = drv. 
```{r}
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv, colour = drv))
```
Now, we will group the categories using the group aesthetic. 

```{r}
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy, group = drv))
```
All the lines are the same colour (blue), but we can change the colour based on drv value. 
```{r}
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy, colour = drv), 
              show.legend = FALSE,)
```
We will now increase complexity, by plotting multiple geoms on a single plot, by adding them together. 

```{r}
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy), colour = "black") +
  geom_point(mapping = aes(x = displ, y = hwy, colour = drv))
```
In the above code, we had to duplicate the line, to save time, we can do the below code. 

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point() +
  geom_smooth()
```
Use mappings in specific layers to enable the display of different aesthetics in different layers. 
This is not styled by class, but you can style the points by themselves... 
So, this makes the points coloured. 

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(colour = class)) +
  geom_smooth()
```
The same goes if you want to specify different data for each layer. We will now use a filter (class = "subcompact") to select a subject of the data and plot only that subset. 

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(colour = class)) +
  geom_smooth(data = filter(mpg, class == "subcompact"), se = FALSE)
```
This was a test to our knowledge, to see what graphs would be made using both the below code. 
It creates the same graph. 
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
geom_point() +
geom_smooth()

ggplot() +
geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) +
geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
```

# 2.22.1 Transformations and Stats - Using Diamonds 

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut))
```
You can see that we didn't add count into the variables, because there is no variable for count. But, R will count all the entries by itself, via geom_bar. Although, we can use stat_count interchangeably with geom_bar - it does the same thing. 
The main reson for this is because every geom has a default stat and every stat has a default geom. 
```{r}
ggplot(data = diamonds) + 
  stat_count(mapping = aes(x = cut))
```

# 2.22.2 Overriding Defaults in ggplot2 

```{r}
demo <- tribble(
  ~cut, ~freq,
  "Fair", 1610, 
  "Good", 4906,
  "Very Good", 12082, 
  "Premium", 13791, 
  "Ideal", 21551)
demo 

ggplot(data = demo) + 
  geom_bar(mapping = aes(x = cut, y = freq), stat = "identity")
```
Here we are overriding mapping from transformed variables to aesthetics. 
```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = after_stat(prop), group = 1))
```

# 2.22.3 Plotting Statistic Details 

This is to show a little more about the transformed variables to aesthetics. 
E.g. we can display a bar chart of the proportion of the total diamond dataset rather than count. 

```{r}
ggplot(data = diamonds) + 
  stat_summary(
    mapping = aes(x = cut, y = depth), 
    fun.min = min, 
    fun.max = max, 
    fun = median
  )
```

# 2.23 Aesthic Adjustments 
Changing aesthetics like colour and fill in the charts! 

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, colour = cut))
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = cut))
```
Now using aesthetics to colour by another variable... clarity. Notice how stacking is done automatically. 
This is done behind the scenes with a position argument. 

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity))
```
The ability to make position adjustments is vital, it allows to customise plots in three ways, identity (raw data), fill (changes heights), and dodge (which forces ggplot2 to not put things on top of each other). 

Using 'position = "identity"' enable us to place each object exactly where it falls in the context of the graph. It is vital for point charts like scatter plots but makes a mess in a bar plot situation. 

```{r}
#To alter transperency (alpha)
ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity)) + 
  geom_bar(alpha = 1/5, position = "identity")
```

```{r}
# To colour the bar outlines with no fill colour. 
ggplot(data = diamonds, mapping = aes(x = cut, colour = clarity)) +
  geom_bar(fill = NA, position = "identity")
```
The position "fill" works like stacking, but makes each set of stacked bars the same height. 

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "fill") 
```
The position "dodge" places overlapping objects directly beside one another. 

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "dodge") 
```
The position "jitter" slightly moves points so you can see them all. 

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "jitter") 
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), position = "jitter") 
```

# 2.24 The Layered Grammer of Graphics 
Template for making a ggplot. 

```{r}
#ggplot(data = <DATA>) + 
#  <GEOM_FUNCTION>(
#    mapping = aes(<MAPPINGS>), 
#    stat = <STAT>, 
#    position = <POSITION>
#  ) + 
#  <FACET_FUNCTION> 
```

#### WORKSHOP 3 ##### 

# 3.1.1 Learning about for-loops 

R & all programming languages have a range of 'control structures that allow you to control the flow of execution of code. 
A primitive way of doing this is with loops. Repeat a task on your data. Code looks like... 

"for (value in that) {
this 
}"

```{r}
for (value in c("My", "first", "for", "loop")){
  print("one run")
}
```
The above printed what we put in the brackets after print i.e. "one run". 

```{r}
for (value in c("My", "second", "for", "loop")){
  print(value)
}
```
But, when we put any word, for example 'value' in the top line, when we ask to print 'value', R will print the words on the first line, as we have asked. Examples are below and above. 

```{r}
for (word in c("My", "second", "for", "loop")){
  print(word)
}
```
We only get an output from our loop because of the 'print' function. If you look to the environment box (top right) you will see the words 'value' and 'word' are loops. 

We will now make our own fully custom for-loop. 

First - we make the object we want to loop through. Here we use a numeric vector, but it will also work with a list (we'll call it a 'list' because it helps imagine it). Print it so you know what it is: note that it's a numeric vector data structure that holds 50 numbers, which is every number from 1 to 50. 

```{r}
my_list <- 1:50
print(my_list) 
str(my_list)
```
str() function compactly displays the internal structure of an object. 

Now, we make a new for-loop for a custom purpose. We can think about it as "for each item on my list, do something to that item". 
Here we want to access the first item, print it out, and then go on to the next item. 

i is what programmers normally use for this purpose (standing for iteration), but you can use any letter or term you like, as long as you call it again inside the for-loop. 

The code will enter into a loop (defined by the curly brackets) then run round and round the loop for however many times your for-loop is looping over. This one runs 50 times, goes to number 1, prints it, goes to number 2, prints it, and so on... 

```{r}
#make a for-loop 
for(i in my_list){
  result <- i 
  out <- paste0("I am printing loop number: ",i,".") #output: 1 loop 
  print(out)
  #Sys.sleep(0.5) #slow it down for example purposes
}

print("For loop complete!")
print("Program complete!")
```
Now we increase the complexity a little. Here we want to add a number constant to every item in the list

```{r}
add_me <- 500
for(i in my_list){
  result <- i + add_me #add to the i value 
  out <- paste0("Result of iteration: ",i,"=", result, ".")
  print(out)
  #Sys.sleep(0.5) #comment out to show just how fast R is. 
}

print("For loop complete!")
print("Program complete!")
```
Now, imagine if you had something very complex to do, like take a data file, extract a value from it, add it to another one and so on. Using this type of functionality of programming language, iterating with little or no input, will ultimately allow you to do incredible things to your data or files. 

Nested for-loop. 
```{r}
#Note we have i and j for each loop 
for(i in c(1, 2, 3, 4, 5)) {
  print (paste ("I am base loop number: ",i))
  #Sys.sleep(0.8) #slow it down for example purposes 
  for (j in c(1, 2, 3)){
    print (paste ("...I am inside loop number: ",j))
    #Sys.sleep(0.8) # slow it down for example purposes
  }
}
```
I will now make my own for-loop. 

```{r}
my_animal_list <- 1:3
print(my_animal_list) 
```

```{r}
animals <- c("cat", "whale", "shark")
```

```{r}
for(a in my_animal_list){
  out <- paste("My number",a,"favourite animal is a", animals[a],".")
  print(out)
  }
```

# 3.1.2 if and if else statements 

if (no_milk){
buy_milk
}

First we tell the code that x is 1. Then we say, if x is 1 then x is 2. Similar to the example above, that says is there is no milk, buy milk. So, now R knows that x is 2. 
```{r}
x <- 1 
x #what is x? 
if (x == 1) {
  x <- 2
  x #what is x now? 
}
```
The below checks to see what x is. We're asking does x = 1 or does x = 2, and we get a TRUE or FALSE answer. 

```{r}
x == 1
```
FALSE

```{r}
x == 2
```
TRUE 

Sometimes, we want to do something when we don't meet a certain case. A marine example of this is to format tracking tag data if there is some data received, but skip the file if no data is received (e.g. the animal is underwater). Here we use the "ifelse" statement, which allows you to start a second set of commands if the first condition is not met. 

The code: 

```{r}
#if(animal_data_receieved){
  #1. save statement in log file that the animal has been updated
#} else {
  #1. save statment saying no data receieved
#}
```

In addition, there is an "elseif" statement which allows you to fit in additional conditions between "if" and "else", enabling you to chain together a range of different cases for your program to move through and do certain things. 

Exercise 

```{r}
name <- "Caterina"
if (name == "Caterina") {
  last.name <- "Palma"
}
print(paste(name, last.name))
```

try changing name to Bob, what happens? 

```{r}
name <- "Bob"
if (name == "Caterina") {
  last.name <- "Palma"
}
print(paste(name, last.name))
```

When I change the first name <- Bob, it works. When i change the second name Bob, it doesnt. 

```{r}
name <- "Caterina"
if (name == "Bob") {
  last.name <- "Palma"
}
print(paste(name, last.name))
```

Now we use else... 

```{r}
name <- "Bob"
if (name == "Caterina") {
  last.name <- "Palma"
} else {
  last.name <- "Smith"
}
print(paste(name, last.name))
```
The above is saying, when the name is Caterina, then the last name is Palma, but if the name is anything else then the last name is Smith. 

```{r}
name <- "Jim"
if (name == "Caterina") {
  last.name <- "Palma"
} else if (name == "Bob") {
  last.name <- "Jones"
} else {
  last.name <- "... last name is unknown"
}
print(paste(name, last.name))
```

The above is saying, if Caterina is Palma, and Bob is Jones, then Jim is ...last name unknown. 

# 3.1.3 Functions 

We just got introduced to control structures such as "if else" statements and "for loops". 
But, if we want to do the same thing over and over again in our script, we use a FUNCTION! 
It makes everything shorter and neater, and saves us from writing long code over and over. 

Let's use a classic problem of wanting to convert temperature. To calculate celcuis from fahrenheit, it's a simple equation. 

```{r}
#temp_c <- (temp_F -32)* 5/9
```

If we want to calculate temperature in celcuis of two places, one which is 100F and the other which is 200F, I could write... 

```{r}
temp_C <-(100 - 32) * 5/9 
temp_C <-(200 - 32) * 5/9 
```

The number comes up in our environment box (top right-hand corner). 

BUT, what if we have a data file where we need to repeatedly make this conversion over hundreds or thousands of sites? Having to include the full equation every time makes for a lot of writing, a big chance of accidently including an error, and difficult to read script. 

Here, we can turn this into a function... 

```{r}
fahrenheit_to_celsuis <- function(temp_F) {
  temp_C <- (temp_F - 32) *5/9 
  return(temp_C)
}
```

The function is called "fahrenheit_to_celcuis", and is defined by assigning the output of 'function'. The list of argument names are contained within parentheses. Arguments are the object we want our function to operate on. In this case, it will be a number which represents temperature in fahrenheit (temp_F). 

Next, the 'body' of the function, which is the code that will be executed when the function is run, is contained within curly brackets. The return function gives us back the result. 

```{r}
fahrenheit_to_celsuis <- function(temp_F) {
  #convert F to C
  temp_C <- (temp_F - 32) *5/9 
  return(temp_C)
}
```

Adding in the # to comment. 

We use the function name "fehrenheit_to_celsuis()" putting a number in the brackets, and R provides us with the answer using the equation we put in the function when we made it earlier/above. 

```{r}
fahrenheit_to_celsuis(100) 
```

```{r}
fahrenheit_to_celsuis(200) 
```

As we change the number of fahrenheit in the (), the value of temp_C changes, you can see this in the environment box (top right-hand corner). This saves a lot of space with the script. 
Using functions for repeated tasks can dramatically improve your coding style and lead to readable, easy to understand code. 

Exercise - Write a function that converts Celsuis to Fahrenheit... 

```{r}
celsuis_to_fahrenheit <- function(temp_C) {
  #convert C to F
  temp_F <- (temp_C * 9/5) + 32
  return(temp_F)
}
```

Now to test... 

```{r}
celsuis_to_fahrenheit(40)
```

# 3.1.5 Importing and Saving Data 

You can download your own data into R as a data frame very easily. 
Generally, we prefer data to be in a comma separated file (.csv) rather than excel (.xls) because .csv's are plain text files that are well structured and easy for R to deal with. 

RStudio has drop down menus and methods to import data into an R session. However, we'll hard-code these types of actions using the base function read.csv, so that it's easier to re-run your script in the future. 

Saving a data frame back to a .csv once you've developed it, simply use write.csv. 

# 3.1.6 Which programming language should I learn? Python vs. R 
This is to show that programming languages are very similar. 

```{r}
#source("Class_1.R")
```

# WORKSHOP 4 # 

#4.4 Chattr (Conversational Workspace Engine)

#Installing chat interfaces. 

```{r}
#install.packages("ellmer")
#install.packages("chattr")

library(ellmer)
library(chattr)
```

#Securely storing our API Key 

Run this code in console: 
```{r}
#usethis::edit_r_environ()
```
Then a new file named .Renviron will open. Add the below line (because we used Gemini): 
GEMINI_API_KEY="your-long-api-key-here" - adding my personal API key in ""

Save file and restart R session by Cmd+Shift+0. 

#Connecting Chattr to Your Model 

Run this code in console: 
```{r}
#usethis::edit_r_profile()
```

Again, a new page will open on R, called .Rprofile 
In this page add one line... options(.chattr_chat = ellmer::chat_google_gemini())  
Save file and restart R session by Cmd+Shift+0. 

#Launching and Mapping the User Interface 

To interact with the AI we have chosen, which is Google Gemini, we can run... 
```{r}
#chattr::chattr_app()
```
We added it as a shortcut by... 
1. Tools > Modify Keyboard Shortcuts 
2. Search Open Chat 
3. Click inside "Shortcut" and add own shortcut... in my case i chose ctrl+o. 
4. Click Apply 

Now my chattr interface is open. When I press ctrl+o it opens in bottom right-hand corner. 

Exercise: Prompting for Context 
Question: I asked "What package would I use to create a scatterplot?" 
Answer: library(ggplot2)

#4.3 Conditional Logical - filtering your Data 

Base R: if and else 

```{r}
#Simple example: Checking Sea Surface Temperature (SST) 
sst <- 30.2

if (sst>29.5) {
  print("Warning: Marine heatwave threshold exceeded!")
} else {
  print("SST remains within baseline parameters.")
}
```

#Tidyverse Alternative: Vectorised Conditionals (dplyr) 
Using the data-wrangling package "dplyr" - if_else()

#IF_ELSE 

The if_else (condition, true_output, false_output) function evaluates an entire column item-by-item. 

```{r}
#Load tidyverse (if not already loaded) 
library(tidyverse)

#Sample coral data 
coral_monitoring <- tibble(
  site = c("Site_A", "Site B", "Site C"), 
  depth_m = c(12, 35, 8)
)
```

```{r}
#Classify depth using if_else 
coral_monitoring <- coral_monitoring %>%
  mutate(zone = if_else (depth_m > 30, "Deep Reef", "Shallow Reef"))
```

What does the function "mutate()" do? 

```{r}
?mutate
```

Mutate creates new columns that are functions of existing variables. It can also modify (if the name is the same as an existing column) and delete columns (by setting their value to NULL). 

What is the %>% symbol? 

```{r}
?'%>%'
```
It passes the result on the left into the function on the right.
It pipes an object forward into a function or call expression. 

#CASE_WHEN 
When you have more than two possibilies, stakcing multiple if-else blocks becomes unreadable. case_when() provides a clean, sequential evaluation layout using formula tildes (~): 

```{r}
coral_monitoring <- coral_monitoring %>% 
  mutate(reef_category = case_when(
    depth_m < 10 ~ "Lagoon /Flats", 
    depth_m <= 30 ~ "Crest/Slope", 
    depth_m > 30 ~ "Mesophotic / Deep", 
    TRUE ~ "Unclassified" # Catch-all remainder 
  ))
```


```{r}
library(tibble)
```

TASK: Environmental Stress Classification 


```{r}
# Create tibble and classify environment type
marine_stations <- tibble(
  salinity = c(35, 28, 32, 12)
) %>%
  mutate(
    environment_type = case_when(
      salinity < 15 ~ "Estuarine",
      salinity >= 15 & salinity <= 30 ~ "Brackish",
      salinity > 30 ~ "Marine"
    )
  )

# Print result to verify
print(marine_stations)

# Check structure
str(marine_stations)
```
#4.4 Automation - Iterating Efficiency 
Automation means instructing R to repeat a task across multiple items without manual copy-pasting. 

Base R: for-loops 
A for-loop repeats a code chunk for each element in a designated sequence. 

```{r}
#Classic loop anatomy 
for(year in 2020:2024){
  print(paste("Porcessing climate data for year:", year))
}
```
If we want to track real-world data, loops can iterate over file directories (e.g., reading 12 monthly turtle tracking CSVs sequentially): 

```{r}
#loop across index sequences 
transect_lengths <- c(50, 100, 25, 75)

for (i in seq_along(transect_lengths)){
  print(paste("Transect number", i, "measures", transect_lengths[i], "meters."))
}
```

# Tidyverse Alnernative: Functional Programming (purrr package) 

While for-loops are a core programming construct, they require you to manage index variables (i) and explicitly set up empty "containers" to store results, which often leads to verbose code and hidden bugs. 

map() = returns a flexible List structure 
map_dbl() = returns a vector of decimals/numeric values 
map_chr() = returns a vector of text strings/characters 
map_df() = returns a combines data frame/tibble 

#Comparison Example: Calculating Square Roots of Site Vectors 

Look at how much cleaner functional mapping behaves compared to building a menual container loop: 

Using a standard for-loop: 
```{r}
site_areas <- c(144, 400, 625)
results <- numeric(length(site_areas)) #must build an empty container first 

for(i in seq_along(site_areas)){
  results[i] <- sqrt(site_areas[i])
}
```

Using the purrr equivalent: 
```{r}
library(purrr)
```
```{r}
#Map directly and define your expexted output type explicityly 
mapped_results <- map_dbl(site_areas, sqrt)
print(results)
```
There is no right or wrong answer. You can just choose what works best for you. 

#Iterating Across Complex Groups 
You can combine purrr pipelines seamlessly inside data frames to split analysis across separate groups: 

```{r}
#Iterate a summary function over split lists of data 
iris %>% 
  split(.$Species) %>% 
  map(~summary(.x))
```

TASK: Using chattr (but I am not using chattr, I am using internet copilot)

Example: 
```{r}

fish_counts <- list(
  c(10, 15, 20),
  c(5, 8, 12),
  c(30, 25, 35)
)
```

For-loop Version: 
```{r}
means_loop <- numeric(length(fish_counts))

for (i in seq_along(fish_counts)) {
  means_loop[i] <- mean(fish_counts[[i]])
}
means_loop
```

purrr version 
```{r}
library(purrr)

means_map <- map_dbl(fish_counts, mean)

means_map
```

Results: the purrr version code is much much shorter and more accurate compared to the example and for-loop version. 

# Writing Custom Functions 
Do this when you find yourself copying and pasting the exact same block of code of loop transformations. 

Functional Construction Blueprint 
A function consists of a Name, Arguments (the input variables), and a Body (the execution logic wrapped inside curly brackets). 

```{r}
#Functional Definition 
calculate_coral_mortality <- function(initial_count, surviving_count) {
  #Logic safety switch using our conditional tools! 
  if(initial_count <=0) {
    stop("Initial count must be greater than zero.")
  }
  mortality_rate <- (initial_count - surviving_count) / initial_count
  return(mortality_rate)
}

calculate_coral_mortality(initial_count = 120, surviving_count = 84)
``` 

Exercise: Custom Function Construction Task 

```{r}
convert_temp_c_to_f <- function(celsius) {
  if (celsius < -273.15) {
    stop("Temperature cannot be below absolute zero (-273.15°C).")
  }
  (celsius * 9/5) + 32
}
# Test
convert_temp_c_to_f(25)    # should work
#convert_temp_c_to_f(-300)  # should error
```
25 degrees Celsius works... -300 degrees Celsius does not work. 

#4.6 Advanced Scripting Extension Challenges 
I copy and pasted the instructions into Copilot and copy and pasted the code they gave me, ran it and it seemed to work with no error. 

```{r}
# Load purrr
library(purrr)

# Input vectors
counts <- c(15, 24, 8, 42)
factors <- c(1.2, 0.8, 2.5, 1.1)

# Map across both vectors simultaneously
scaled_abundance <- map2_dbl(counts, factors, ~ .x * .y)

# View result
scaled_abundance
```

```{r}
# Load libraries
library(dplyr)
library(tibble)

# Define the function
clean_and_classify_survey <- function(df, column_name, threshold) {
  
  # Check if column exists
  if (!(column_name %in% names(df))) {
    stop(paste("Error: Column", column_name, "does not exist in the data frame."))
  }
  
  # Clean and classify
  result <- df %>%
    filter(!is.na(.data[[column_name]])) %>%   # remove NAs in specified column
    mutate(
      classification = if_else(
        .data[[column_name]] >= threshold,
        "Above threshold",
        "Below threshold"
      )
    )
  
  # Return result (also assign to global env if needed)
  assign("cleaned_survey", result, envir = .GlobalEnv)
  
  return(result)
}

#Create mock tibble with NA
survey_data <- tibble(
  site = c("A", "B", "C", "D"),
  abundance = c(10, NA, 25, 5)
)

#Test function
clean_and_classify_survey(survey_data, "abundance", 15)
```


